---
title: "Functionalizing Rasters from NEON"
author: "Kristin Braziunas"
date: "June 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

Before starting on the objectives, I need to open the packages I will need.

```{r load-libraries-functions }

library(raster)
library(rgdal)
library(neonAOP)
source("scripts/functions.R")

```

1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)

```{r import-chm-raster }

# import CHM
chm <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")
chosen.breaks <- c(6, 10, 20, 40, 56)  # set parameters

# take a look at it
plot(chm, main = "Canopy Height Model (CHM) \nTEAK field site")

```

2. For the CHM, set values == 0 to NA (not trees)

```{r set-chm-na-values }

# are there 0 values?
hist(chm)  # looks like a lot

# set NA values
chm[chm == 0] <- NA

# make sure it worked with a histogram
hist(chm)  # looks better

# can also look at density plot
density(chm, xlab="Height (m)")

```

3. Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize histogram/density and plot vertical cutoff lines.

```{r classify-raster }

# look at CHM
chm  # max value is 55.68

density_with_breaks(rast.in = chm, title = "TEAK CHM density", bins=chosen.breaks)

# reclassify the raster

chm.htclass <- reclassify(chm, create_height_class_matrix(chosen.breaks))

# look at plot
plot(chm.htclass, main = "Reclassified CHM by height classes \nTEAK field site")

```


## Save our density plot as a PDF

```{r save-pdf-density, eval=FALSE }

# save as pdf using new function
make_pdf(density_with_breaks(rast.in = chm, 
                             title = "TEAK CHM density", 
                             bins=chosen.breaks), 
         filename = "TEAK_CHM_density_with_breaks.pdf")

```

4. Take the chm and create a hillshade (http://neon-workwithdata.github.io/neon-data-institute-2016/R/create-hillshade-R/)

```{r create-hillshade, eval = FALSE }

# following along with NEON website
# first have to import dsm

dsm <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarDSM.tif")

# extract slope and aspect
slope <- terrain(dsm, opt="slope")
aspect <- terrain(dsm, opt="aspect")

# create hillshade
dsm.hill <- hillShade(slope, aspect,
                      angle = 40,
                      direction = 270)

# stopped here, because not creating a hillshade right now

```

5. PLOT - Plot the classified raster, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)

6. Export the plot figure to a pdf – publishable


```{r plot-classified-raster }

plot_reclassified_raster(chm.htclass, site.name = "TEAK", breaks = chosen.breaks)

```

```{r save-pdf-classified-raster, eval=FALSE }

# save as PDF
make_pdf(plot_reclassified_raster(chm.htclass, site.name = "TEAK",
                                  breaks=chosen.breaks),
         filename = "TEAK_chm_reclassified.pdf")

```


7. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.

```{r export-geotiff, eval=FALSE }

writeRaster(chm.htclass,
            file = "../outputs/TEAK/TEAK_CHM_classified.tif",
            format = "GTiff",
            options = "COMPRESS=LZW", 
            overwrite = TRUE,
            NAflag = -9999)

```

```{r timestamp-log }

# create timestamped logfile
sink(paste0(format(Sys.time(), "%Y-%m-%d_%H%M%S"),
            "_sessionInfo.txt"))
sessionInfo()
sink()

```

## Adding make PDF function

```{r add-make-pdf-function }

hist_from_chm <- function(chm, title, breaks) {
  hist(chm, main = title, xlab = "Canopy heights (m)")
  abline(v = breaks, col = "red")
}

hist_from_chm(chm, title = "Histogram of TEAK CHM", breaks = c(10, 30))
make_pdf(hist_from_chm(chm, 
                       title = "Histogram of TEAK CHM", 
                       breaks = c(10, 30)), 
         "TEAK_CHM_histogram.pdf")

```

